---
title: "Logistic models to assess tolerance"
author: "Madyline Lawrence, Moria Chambers & Owais Gilani"
date: "6/17/2022"
editor_options:
  chunk_output_type: console
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 4
    theme: cerulean
    code_folding: hide
---
# Goal: To determine how chronic infection alters the ability to tolerate secondary infections of unrelated bacteria

## Ready the workspace
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Importing packages needed

library(gdata)
library(ggplot2)
library(stringr)
library(boot)
library(data.table)
library(dplyr)
library(formattable)
library(tidyr)
library(caret)
library(questionr)
library(DescTools)
library(survival)
```

```{r}
rm(list=ls()) #removes all old variables

#import the dataset
efdata = read.csv("E.fae data.csv")
smdata = read.csv("S.m data.csv")

ef.bygroup = distinct(efdata, scientist,rep, primary, secondary_dose, X0.hour, X10.hour, log10_10hour)
sm.bygroup = distinct(smdata, scientist, rep, primary, secondary_dose, X0.hour, X10.hour, log10_10hour)
```

# Chronic infection with *S. marcescens* {.tabset}

## Full Logistic Model
  + Does chronic infection with *S. marcescens* significantly impact tolerance to secondary infection with *P. rettgeri*?
  + We assessed this by determining whether chronic infection significantly impacts the relationship between log 10 (bacterial load at 10 hours post-secondary infection) and survival three days post-secondary infection.
```{r}
mylogit = glm(formula = bin_alive3 ~ log10_10hour + bin_primary + (bin_primary*log10_10hour), family = "binomial", data = smdata)
summary(mylogit)

# C-Statistic
cat(paste("C-Statistic =", round(Cstat(mylogit), 4)))
```
  + **Conclusion** - Yes, chronic infection with *S. marcesens* has a significant impact on survival three days post-infection (p = 0.02) and significantly impacts the relationship between log(10 hour bacterial load) and survival at three days post-infection (p << 0.001).

## Values of Interest

```{r}
# Function to calculate Inflection Point
inflection.point <- function(beta0, beta1){
  return(as.numeric(-beta0/beta1))
}

# Function to calculate Hill's Slope
hill.slope <- function(beta0, beta1){
  x <- -beta0/beta1
  u <- beta0+beta1*x
  return(as.numeric(beta1*exp(u)/(1+exp(u))^2))
}
```

```{r eval=FALSE}
# PBS curve point of inflection
inflection.point(beta0=mylogit$coefficients[1], beta1=mylogit$coefficients[2])

# PBS Hill's slope
hill.slope(beta0=mylogit$coefficients[1], beta1=mylogit$coefficients[2])

# S. Marcescens Curve point of inflection
inflection.point(beta0=(mylogit$coefficients[1]+mylogit$coefficients[3]),
                 beta1=(mylogit$coefficients[2]+mylogit$coefficients[4]))

# S. Marcescens Hill's slope
hill.slope(beta0=(mylogit$coefficients[1]+mylogit$coefficients[3]),
           beta1=(mylogit$coefficients[2]+mylogit$coefficients[4]))
```
  + PBS curve point of inflection = 3.6981
  + PBS hill slope = -0.26585
  + *S. marcescens* curve point of inflection = 5.71890
  + *S. marcescens* hill slope = -0.10105

## Figure 3C
```{r}
bact_val <- seq(2,9,0.01)
PBS_predict_data <- data.frame(log10_10hour = bact_val, 
                               bin_primary = c(rep(0, length(bact_val))))
# PBS Prediction
PBS_preds <- predict(mylogit, newdata = PBS_predict_data, type = "response", se.fit = TRUE)
PBS_upper = (PBS_preds$fit+(1.96*PBS_preds$se.fit))
PBS_lower = (PBS_preds$fit-(1.96*PBS_preds$se.fit))

# SM Prediction
SM_predict_data <- data.frame(log10_10hour = bact_val, 
                              bin_primary = c(rep(1, length(bact_val))))
SM_preds <- predict(mylogit, newdata=SM_predict_data, type = "response", se.fit = TRUE)
SM_upper = (SM_preds$fit+(1.96*SM_preds$se.fit))
SM_lower = (SM_preds$fit-(1.96*SM_preds$se.fit))
```

```{r, echo = FALSE}
#png(file="Figure3C.png", width = 1500, height = 1500, res = 300)
#par(family = "Times")
plot(0,0, type = "n",
    bty = "n",
    lwd=2, #line width
    yaxt='n',
    xaxt='n',
    ylab="probability of survival", #label for the y-axis
    xlab=expression("log"[10]*"(bacterial load at 10 hours post-infection)"), #label for the x-axis
    cex.lab=1.3, #font size of axis labels
    xlim = c(2, 9), #determines range of the x-axis
    xaxs='i', #makes the data flush with the axis 
    ylim = c(0, 1), #determines range of the y-axis
    yaxs='i') #makes the data flush with the axis
axis(2, #left side
     las=2, #perpendicular to the axis
     cex.axis=1, #font size
     lwd=2) #line width
axis(1, #bottom of the graph
     cex.axis=1, #font size
     lwd=2) #line width

polygon(c(bact_val, rev(bact_val)), c(PBS_upper, rev(PBS_lower)), border = NA, col = adjustcolor("gray", alpha.f=0.6))
polygon(c(bact_val, rev(bact_val)), c(SM_upper, rev(SM_lower)), border = NA, col = adjustcolor("#055df5", alpha.f=0.6))
lines(PBS_preds$fit~bact_val,col="gray",lty=2, lwd = 3)
lines(SM_preds$fit~bact_val, col="#055df5",lty=1, lwd = 3)

legend("topright", c("chronically infected", "control"), col=c("#055df5","gray"),lty = c(1,2), lwd = 3, cex = 1.3, bty = "n")
lines(bact_val, PBS_upper, lty=1, lwd = 1, col=adjustcolor("gray", alpha.f=0.7))
lines(bact_val, PBS_lower, lty=1, lwd = 1, col=adjustcolor("gray", alpha.f=0.7))
lines(bact_val, SM_upper, lty=1, lwd = 1, col=adjustcolor("#055df5", alpha.f=0.7))
lines(bact_val, SM_lower, lty=1, lwd = 1, col=adjustcolor("#055df5", alpha.f=0.7))

#dev.off()
```


# Chronic infection with *E. faecalis* {.tabset}

## Full Logistic Model
  + Does chronic infection with *E. faecalis* significantly impact tolerance to secondary infection with *P. rettgeri*?
  + We assessed this by determining whether chronic infection significantly impacts the relationship between log 10 (bacterial load at 10 hours post-secondary infeciton) and survival three days post-secondary infection.
```{r}
mylogit_Ef = glm(formula = bin_alive3 ~ log10_10hour + bin_primary + (bin_primary*log10_10hour), family = "binomial", data = efdata)
summary(mylogit_Ef)

# C-Statistic
cat(paste("C-Statistic =", round(Cstat(mylogit_Ef), 4)))
```
  + **Conclusion** - No, chronic infection with *E. faecalis* does not significantly impact the relationship between log(10 hour bacterial load) and survival at three days post-infection (p = 0.52). Therefore we removed this from the model to better assess the overall impact of chronic infection with *E. faecalis*

## Reduced Logistic Model

```{r}
## Since the interaction term isn't significant, removed it from the model and reran
mylogit_Ef_redux = glm(formula = bin_alive3 ~ log10_10hour + bin_primary, family = "binomial", data = efdata)
summary(mylogit_Ef_redux)
anova(mylogit_Ef_redux, mylogit_Ef)

cat(paste("C-Statistic =", round(Cstat(mylogit_Ef_redux), 4)))
```
  + **Conclusion** - Chronic infection with *E. faecalis* significantly impacts survival at three days post-infection (p<0.001)

## Values of Interest

```{r}
# Function to calculate Inflection Point
inflection.point <- function(beta0, beta1){
  return(as.numeric(-beta0/beta1))
}

# Function to calculate Hill's Slope
hill.slope <- function(beta0, beta1){
  x <- -beta0/beta1
  u <- beta0+beta1*x
  return(as.numeric(beta1*exp(u)/(1+exp(u))^2))
}
```

```{r eval=FALSE}
# PBS curve point of inflection
inflection.point(beta0=mylogit_Ef_redux$coefficients[1], beta1=mylogit_Ef_redux$coefficients[2])

# PBS Hill's slope
hill.slope(beta0=mylogit_Ef_redux$coefficients[1], beta1=mylogit_Ef_redux$coefficients[2])

# E.faecalis Curve point of inflection
inflection.point(beta0=(mylogit_Ef_redux$coefficients[1]+mylogit_Ef_redux$coefficients[3]),
                 beta1=(mylogit_Ef_redux$coefficients[2]))

# E.faecalis Hill's slope
hill.slope(beta0=(mylogit_Ef_redux$coefficients[1]+mylogit_Ef_redux$coefficients[3]),
           beta1=(mylogit_Ef_redux$coefficients[2]))
```
  + PBS curve point of inflection = 3.55631 
  + PBS hill slope = -0.14356
  + *E.faecalis* curve point of inflection = 4.55966
  + *E.faecalis* hill slope = -0.14356


## Figure 3D
```{r}
bact_val = seq(2,9,0.01)
PBS_predict_data <- data.frame(log10_10hour = bact_val, 
                               bin_primary = c(rep(0, length(bact_val))))

PBS_preds <- predict(mylogit_Ef_redux, newdata=PBS_predict_data, type = "response",se.fit =TRUE)
PBS_upper = (PBS_preds$fit+1.96*PBS_preds$se.fit)
PBS_lower = (PBS_preds$fit-1.96*PBS_preds$se.fit)

Efae_predict_data <- data.frame(log10_10hour = bact_val, 
                                bin_primary = c(rep(1, length(bact_val))))

Efae_preds <- predict(mylogit_Ef_redux, newdata=Efae_predict_data, type = "response",se.fit = TRUE)
Efae_upper = (Efae_preds$fit+(1.96*Efae_preds$se.fit))
Efae_lower = (Efae_preds$fit-(1.96*Efae_preds$se.fit))
```

```{r, echo = FALSE}
#png(file="Figure3D.png", width = 1500, height = 1500, res = 300)
#par(family = "Times")
plot(0,0, type = "n",
    bty = "n",
    lwd=2, #line width
    yaxt='n',
    xaxt='n',
    ylab="probability of survival", #label for the y-axis
    xlab=expression("log"[10]*"(bacterial load 10 hours post-infection)"), #label for the x-axis
    cex.lab=1.3, #font size of axis labels
    xlim = c(2, 9), #determines range of the x-axis
    xaxs='i', #makes the data flush with the axis 
    ylim = c(0, 1), #determines range of the y-axis
    yaxs='i') #makes the data flush with the axis
axis(2, #left side
     las=2, #perpendicular to the axis
     cex.axis=1, #font size
     lwd=2) #line width
axis(1, #bottom of the graph
     cex.axis=1, #font size
     lwd=2) #line width

polygon(c(bact_val, rev(bact_val)), c(PBS_upper, rev(PBS_lower)), border = NA, col = adjustcolor("gray", alpha.f=0.6))
polygon(c(bact_val, rev(bact_val)), c(Efae_upper, rev(Efae_lower)), border = NA, col = adjustcolor("#055df5", alpha.f=0.6))
lines(PBS_preds$fit~bact_val,col="gray",lty = 2, lwd = 3)
lines(Efae_preds$fit~bact_val, col="#055df5",lty = 1, lwd = 3)

legend("topright", c("chronically infected", "control"), col=c("#055df5","gray"), lty=c(1,2), lwd = 3, cex = 1.3, bty = "n")
lines(bact_val, PBS_upper, lty=1, lwd = 1, col=adjustcolor("gray", alpha.f=0.7))
lines(bact_val, PBS_lower, lty=1, lwd = 1, col=adjustcolor("gray", alpha.f=0.7))
lines(bact_val, Efae_upper, lty=1, lwd = 1, col=adjustcolor("#055df5", alpha.f=0.7))
lines(bact_val, Efae_lower, lty=1, lwd = 1, col=adjustcolor("#055df5", alpha.f=0.7))
#dev.off()
```
